\documentclass[12pt]{article}
\usepackage{../../template}
\title{Lecture 28}
\author{niceguy}
\begin{document}
\maketitle

\section{Extended Integrals}

$f: A \rightarrow \R$ is continuous, with $A$ being open. Then $f$ is extended integrable iff
$$\sup_{D\in A} \int_D |f| < \infty$$
where $D$ is compact and rectifiable. Then
$$\int_A f = \int_ f_+ - \int_A f_-$$

$\forall A \subseteq \R^n$ open $\exists$ a sequence $C_n \subset A$ such that $C_n \subset \text{ Int} C_{n+1}, C_n$ is compact and rectifiable.

\begin{prop}
    $f(x)$ is extended integrable over A iff
    $$\lim_{n\rightarrow\infty} \int_{C_n} |f| < \infty$$
    Moreover,
    $$\int_A f = \lim_{n\rightarrow\infty} \int_{C_n} f_+ - \lim_{n\rightarrow\infty} \int_{C_n} f_-$$
\end{prop}

\begin{proof}
    First assume $f$ is non-negative, so $f = |f|$. Then $\int_{C_n} f$ increases; it converges iff it is bounded. Suppose again $f$ is integrable over $A$. If we denote compact and rectifiable subsets of $A$ by $D$, then
    $$\int_{C_n} f \leq \sup_D \int_D f = \int_A f$$
    Then the sequence has an upper bound of $\int_A f$, so it converges. Conversely, if the sequence is bounded/converges, let $D$ be a compact and rectifiable subset of $A$. It is covered by the countable sequence of $C_n$. Since it is compact, it is covered by finitely many of them. Since each is a subset of the next, it is covered by one, or the interior of the next one. Then
    $$\int_D f \leq \int_{C_m} f \leq \lim_{n\rightarrow\infty} \int_{C_n} f$$
    So $\int_D f$ is bounded above for arbitrary $D$. By definition of the extended integral, $\int_A f$ exists. \\
    Note that by definition, $0 \leq f_+(x) \leq |f(x)|, 0 \leq f_-(x) \leq |f(x)|$, and $|f(x)| = f_+(x) + f_-(x)$. By definition, $f$ is integrable over $A$ iff both $f_+,f_-$ are integrable over $A$, which occurs iff $f_+,f_-$ integrated over $C_n$ is bounded. From the above, they are bounded iff the integral of $|f|$ over $C_n$ is bounded. Then
    \begin{align*}
        \lim_{n\rightarrow\infty} \int_{C_n} f &= \lim_{n\rightarrow\infty} \int_{C_n} f_+ - \lim_{n\rightarrow\infty} \int_{C_n} f_- \\
                                               &= \int_A f_+ - \int_A f_- \\
                                               &= \int_A f
    \end{align*}
\end{proof}

\end{document}
