Here is a bunch of stuff that you have to do, and some stuff that you should.

\section{Necessary}

\subsection{Study Permit}

You need a study permit. Apply as soon as possible. Check out more at \url{https://www.canada.ca/en/immigration-refugees-citizenship/services/study-canada/study-permit.html}.

\subsection{Winter Wear}

Ignore this if you come from someplace cold. This is for those of you used to 10 degree winters. Get winter boots, but you can use hiking boots if you have them. IMO it is more important for footwear to prevent you from slipping than it is to keep your feet warm, because it isn't that cold in Toronto, and you don't spend a lot of time outdoors in winter.

As for winter clothing, dress up in layers! Check out for sales, especially in the summer. Get more than you feel you need for your first winter; better be safe than sorry.

\section{Optional}

\subsection{International Fee Exemption}

In general, if you are the dependent of a Canadian citizen, permanent resident (approved "in-principle" also counts), or someone with a work permit, you \textbf{may} qualify for an international fee exmeption. You're going to save a lot of money, so please do.

\subsection{Lester B. Pearson Scholarship}

From the website, "[t]he Lester B. Pearson International Scholarships will cover tuition, books, incidental fees, and full residence support for four years." Worth a try to apply. \href{https://future.utoronto.ca/pearson/about/}{More information}.

\subsection{Language Skills}

I'll actually be serious in this part. We have people from all around the world here in U of T, and you should never be ashamed of speaking or writing because you feel your English is not good. Having gone through it myself, I understand the struggles of learning English as a second language. People are not going to nitpick on your grammar or pronunciation. You'll also get loads of opportunities to practice English. To be honest, I would be more concerned with preserving my first language than learning English here in Canada.

\subsection{Cultural Differences}

Depends on where you come from, 一言難盡.
