\begin{enumerate}
    \item You should therefore spend a total of about 8.5 to 11 hours per week on this course. This matches the expectation that full-time Engineering Science students with 6 courses will spend about 50 to 66 hours per week on their studies.
    \item This needs to be said more but your marks don't reflect who you are or your capabilities, with the exception of your Praxis marks. Your marks in Praxis I II and III are very indicative of your abilities as a student engineer as well as your value as a human being. My advice would be to focus on maintaining a 95+ Praxis average if you want to be even slightly successful and respected.
    \item there are 2 stages in life. When you're in highschool, you say "i can't believe it's only a tuesday or a wednesday. it totally feels like a friday... why can't it be the weekend already". When you're in eng sci, you say "thank FUCK it's a tuesday i can't get half this shit done by next monday, every sleeping hour is a resource which i must abuse"
    \item If I don't get this pendulum done by Friday, I'll hang myself at the praxis lecture room and BE the pendulum
    \item bro just like how home is where ur family is, Praxis is where Bobbert is, if there is no Bobbert, then it ain't Praxis to me
    \item me when boofmeister is doing the nerve impulse demonstration and i run down and hook him up to a car battery
    \item perhaps if you had a fried egg and rice it would be easier for you to apply reflexive principalism
    \item Don't worry, it is never too late to switch programs!
    \item When there are more than 50 messages in a channel, I know something is due
\end{enumerate}
