\documentclass[12pt]{article}
\usepackage{../../template}
\author{niceguy}
\title{Lecture 16}
\begin{document}
\maketitle

\section{Recap}

The Poission distribution is
$$p(x;\lambda) = \frac{e^{-\lambda}\lambda^x}{x!}$$
The exponential distribution is
$$f(x;\beta) = \frac{1}{\beta}e^{-\frac{x}{\beta}}$$
Where $\beta = \frac{1}{r}$, the time between events.

\begin{ex}
	In a factory, components fail every 4 days on average. What is the chance a component lasts longer than a week? Note that the exponential distribution is used. \\
	We have $\beta = 4$. Then
	$$P(X\geq7) = \int_7^\infty \frac{1}{4}e^{-\frac{x}{4}}dx = e^{-\frac{7}{4}} \approx 0.17$$
	Note that we expect $\frac{7}{4}$ failures in one week.
\end{ex}

\section{Memoryless Proposition}

Given a random variable with exponential distribution,
$$P(X\geq s+t|X\geq s)$$

By definition of conditional probability, this is equal to
$$\frac{P(X\geq s+t)}{P(X\geq s)}$$
Now
$$P(X\geq s) = \int_s^\infty \frac{1}{\beta}e^{-\frac{x}{\beta}}dx = e^{-\frac{s}{\beta}}$$
Then the quotient is $e^{-\frac{t}{\beta}}$, or $P(X\geq t)$.

\section{Functions of Random Variables}

For the discrete case, consider $X$ with distribution $f(x)$. Let $Y=u(X)$, where $u$ is bijective. Then we can write $X = u^{-1}(Y)$. Let $g(y)$ the the distribution of $y$.
\begin{align*}
	g(y) &= P(Y=y) \\
	     &= P(u^{-1}(Y)=u^{-1}y) \\
	     &= P(X=u^{-1}y) \\
	     &= P(X=u^{-1}y) \\
	     &= f(u^{-1}(y))
\end{align*}

For the continuous case, let $G(y)$ be the cumulative distribution. Then
\begin{align*}
	G(y) &= P(Y\leq y) \\
	     &= P(u^{-1}(Y) \leq u^{-1}(y)) \\
	     &= P(X\leq u^{-1}(y)) \\
\end{align*}

Then
\begin{align*}
	g(y) &= \frac{d}{dy}G(y) \\
	     &= \frac{d}{dy}\int_{-\infty}^{u^{-1}(y)} f(t)dt \\
	     &= f(u^{-1}(y))\frac{du^{-1}(y)}{dy}
\end{align*}

$u$ is either strictly increasing or strictly decreasing. In the latter case, we have an extra negative sign in the slope, giving us
$$g(y) = f(u^{-1}(y))\Bigg |\frac{du^{-1}(y)}{dy} \Bigg |$$
\end{document}
