\documentclass[12pt]{article}
\usepackage{../../template}
\title{Lecture 1}
\author{niceguy}
\begin{document}
\maketitle

\begin{defn}
 A differential equation is an equation that relates a function with its derivatives.
\end{defn}

\begin{ex}
 ODEs:
 
\begin{itemize}
 \item $y'(t) = y(t)$ \\
 \item $y''(t) = 6y'(t) + 2y(t) + 8$ \\
 \item $y'''(t) = [y'(t)]^2$ \\
\end{itemize}

\begin{itemize}
 \item $u_t(t,x,y,z) = u_{xx} + u_{yy} + u_{zz}$ \\
\end{itemize}
\end{ex}

Differential equations can be used to model physical phenomenons, such as weather forecasting, pricing financial derivatives, population dynamics, etc.

\section{Newton's Law of Cooling}

Form a \textit{simple} model for the temperature of a cup of coffee sitting in a room. We let $u(t)$ be the temperature of coffee at time $t$, and $T$ be the (constant) ambient temperature. \\
First model:
$$\frac{du}{dt} = -(u(t) - T)$$

where the temperature of the cup tends to the ambient temperature at a rate proportional to the difference in temperatures. \\
Second model:
$$\frac{du}{dt} = -k(u(t) - T), k > 0$$

where $k$ is a transmission coefficient, representing factors such as the thermal conductivity of the cup. \\

\subsection{Question of Interest}
\begin{enumerate}
 \item Existence of a solution \\
 \item Uniqueness \\
\end{enumerate}

First assume $u \neq T$. Then division gives us
\begin{align*}
 \frac{1}{u - T} \frac{du}{dt} &= -k \\
 \frac{1}{u - T} du &= -kdt \\
 \ln |u - T| &= -kt + C \\
 u - T &= \pm e^Ce^{-kt} \\
\end{align*}

Hence $u(t) = \tilde{C}e^{-kt} + T, \tilde{C} \in \mathbb{R} \backslash \{0\}$ \\

If $u = T$, then there is an equilibrium. \\
All solutions can be written in the general form of
$$u(t) = \tilde{C}e^{-kt} + T, \tilde{C} \in \mathbb{R}$$

To verify this, we can differentiate $u(t)$ as given by the solution
 $$u'(t) = -\tilde{C}ke^{-kt} = -k(u(t) - T)$$
 
Uniqueness: \\

Giving just an ODE will not be enough for a unique solution. Initial conditions must be given. \\

Question: given a differential equation and initial conditions is it easy to see if the solution will be unique or even exist? \\
Answer: No. \\
\end{document}  
